---
title: "Model details"
author: 
- email: s.contractor@unsw.edu.au
  name: It's academic team
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2
fontsize: 12pt
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
bibliography: "https://api.citedrive.com/bib/bbf809b5-e0a6-41ec-a8e1-4e055cef8f9f/references.bib?x=eyJpZCI6ICJiYmY4MDliNS1lMGE2LTQxZWMtYThlMS00ZTA1NWNlZjhmOWYiLCAidXNlciI6ICI2NTEwIiwgInNpZ25hdHVyZSI6ICIyMTIyZDlkODFkNWI5MjFiYWYzOGU1MDExMzUwYjUxMTNiMDUxYmQxYzhkZGE5Y2M4YjM3YjdlZjViYzA5MDgyIn0=/bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F)
```

This document complements information presented in the dashboard. It contains technical information about the statistical modelling including the purposes of the model, associated statistical methodology and additional model results and validation diagnostics. Specifically, visualization and diagnostics associated with the predictive performances of the spatial Bayesian models. A leave-out data analysis is performed to assess model predictions with the exclusion of 20% of data randomly for each city using five prediction-performance measures.   

### Spatial model {-}
The spatial statistical model is developed to predict temperature anomalies (relative to 1981-2000 mean temperatures) across 5x5 km gridded locations within city bounding boxes. The model uses spatial random effects that are based on an automatic specification of spatial basis functions as well as fixed effects including greenness index, human population density and urban area. The model is fitted under an hybrid Bayesian and Frequentist framework using the R package FRK version 2.1.5 [@sainsbury2021modelling] and a normal probability distribution. Statistical and computational details can be found in the two main FRK papers  [@Zammit_Mangion_2021; @sainsbury2021modelling]. Model codes are available at <https://github.com/UN-Datathon-Its-Academic> for reproducibility. 

```{r, echo = FALSE}
rm(list=ls())
source("R/packages.R")
source("R/functions.R")

#Import all data from the Data folder 
datalist <- list.files(path = "Data/", pattern='_merged.csv$', all.files= T, full.names= T)

# Predictive performance 
list.indicators <- list()
pred_sum_list <- list()
```

```{r, include=FALSE}
for ( i in 1:length(datalist)){

# Full dataset 
 X_orign <- read.csv(datalist[i]) %>%
    mutate(City = str_match(datalist[i], "/\\s*(.*?)\\s*_")[2])

  # Select 80%
 X <- X_orign %>% sample_frac(.80)

 ## Convert to sf
  grid_pred <- st_as_sfc(X$geometry)
  st_geometry(X) <- grid_pred
  
# Remaining 20%
  test  <- anti_join(X_orign, X , by = 'system.index')
  grid_pred_test <- st_as_sfc(test$geometry)
  st_geometry(test) <- grid_pred_test

   ## Convert to sf
  grid_pred <- st_as_sfc(X_orign$geometry)
  st_geometry(X_orign) <- grid_pred

# Making STObj - response variable 
  X_Obj <- X %>% dplyr::select(temperature_anomaly) 
  
# Making BAUs - other covariates
   ST_BAUs <- as_Spatial(X_orign %>% dplyr::select(!temperature_anomaly)%>% 
                          mutate(dplyr::across(c(building_count:greenness), ~ as.numeric(scale(.)))))
  
  
  ST_BAUs$fs <- 1                   # scalar fine-scale variance matrix

  # Create basis functions
  basis <- auto_basis(manifold = plane(),      # functions on sphere
                      data=ST_BAUs, 
                      nres = 3,                # use three DGGRID resolutions
                      prune= 0                 # prune basis functions
  )  
  
  ### Model 
  mod <- FRK(f = temperature_anomaly ~ 1 + greenness + population_density + total_area, 
             data = list(as_Spatial(X_Obj)), 
             BAUs = ST_BAUs, 
             basis = basis, 
             response = "gaussian", 
             K_type = "precision", 
             method = "TMB", 
             est_error = TRUE) %>%
          suppressMessages() %>%
          suppressWarnings()

  pred <- predict(mod, type = c("mean"))

  # Extracting posterior distributions of predictive locations 
  
  post_dist_df <- as.data.frame(pred$MC$mu_samples) %>% 
    mutate(system.index = ST_BAUs@data$system.index) %>%
    tidyr::pivot_longer(!system.index,
                        names_to = "draw", 
                        values_to = "pred"
    )
  
  test_pred <- inner_join(test, post_dist_df, by = "system.index")

  test_pred_sum <- test_pred %>% group_by(system.index) %>% 
  ggdist::median_hdci(pred)%>%
  inner_join(test_pred %>% group_by(system.index) %>% slice(1) %>%
               dplyr::select(system.index,temperature_anomaly)) %>%
    mutate(p_se = .upper - .lower) %>%
    mutate(City = str_match(datalist[i], "/\\s*(.*?)\\s*_")[2])
  
pred_sum_list[[i]] <- test_pred_sum 

## Measures

list.indicators[[i]]<- c(test_pred_sum$City[1],round(coverage95(test_pred_sum$temperature_anomaly, test_pred_sum$.lower, test_pred_sum$.upper),2),
  round(IS95(test_pred_sum$temperature_anomaly, test_pred_sum$.lower, test_pred_sum$.upper),2),
  round(RMSPE(test_pred_sum$temperature_anomaly, test_pred_sum$pred),2),
  round(crps(test_pred_sum$temperature_anomaly, data.frame(test_pred_sum$pred, test_pred_sum$p_se))$CRPS,2), round(AIC(mod),2))

}
```

### Fixed effects {-}

```{r, echo = FALSE}
mod_list <- readRDS("Data/models.rds")
coef_table_list <- list()

for ( i in 1:length(mod_list)){
coef_table_list[[i]] <- coef_uncertainty(mod_list[[i]], percentiles = c(2.5, 50, 97.5), nsim = 400, random_effects = FALSE)%>%
  data.frame() %>%
  tibble::rownames_to_column() %>%
  tidyr::pivot_longer(cols =  !rowname, names_to = "param", values_to = "value")%>%
  tidyr::pivot_wider(names_from = rowname, values_from = value) %>%
  mutate(City =  mod_list[[i]]@BAUs$City[1])
}

coef_table <- data.frame(do.call(rbind,coef_table_list))
#coef_table$param <- c("Intercept", "Greenness", "Human population density", "Building area")

coef_table <- coef_table %>%
  mutate(param = case_when(param == "greenness" ~ "Greenness", 
                           param == "population_density" ~ "Population density",
                           param == "total_area" ~ "Urban area",
                           param == "Intercept" ~ "Intercept"))
  
p_coef <- ggplot(coef_table %>% filter(!param == "Intercept"), aes(y=param, x=X50.))+ geom_point() +
  geom_errorbar(aes(y = param, xmin = X2.5., xmax = X97.5.), width=.1) + facet_wrap(~ City) +
  geom_vline(xintercept = 0, linetype = "dashed") +theme_bw() +
  xlab("Effect size") + ylab("")
```

```{r fig1, fig.align = 'center', fig.width=7,fig.height=7, fig.cap="Effect size of the model covariates for each city. Dots shows averaged size effect and error bars corresponding to the 95% credible intervals.", echo=FALSE}
p_coef
```


```{r, echo = FALSE}
pred_table <- data.frame(do.call(rbind,pred_sum_list))
indicators_table <- data.frame(do.call(rbind,list.indicators))
```

```{r, echo = FALSE}
p.pred <- ggplot(pred_table, aes(x=temperature_anomaly, y=pred)) + 
   geom_abline(col="green", linetype = "dashed")+ 
   geom_point() + facet_wrap(~City, scales = "free", ncol = 3) + 
   theme_bw()+ xlab("Observed temperature anomaly") + ylab("Predicted temperature anomaly") + 
   theme(axis.text.x = element_text(size=11),
       axis.text.y = element_text(size=11),axis.title.y=element_text(size=12),
       axis.title.x=element_text(size=12))
```
### Model predictive capability {-}

```{r fig2, fig.align = 'center', fig.width=6,fig.height=7, fig.cap="Predicted temperature anomalies (y axis) relative to true temperature anomalies (x axis).", echo=FALSE}
p.pred
```

With the exception of Singapore which had only 23 data points to train on, our predictions lie very close to the 1-1 line.

### Model diagnostics {-}

```{r, echo = FALSE}
knitr::kable(data.frame(indicators_table), "pipe", 
             col.name = c("City","95% coverage", "95% interval score", "Root-mean-squared prediction error", "Continuous ranked probability score", "AIC"),
             align = c("l","c","c","c","c"), caption = "Measures of performance for model forecasts")
```
